---
title: "Interpretable Machine Learning"
output: 
  html_notebook:
    toc: yes
    toc_depth: 4
    toc_float: true
    theme: flatly
    highlight: zenburn
---

```{r setup, include=FALSE}
library(rmarkdown)
library(knitr)
library(dplyr)
library(ggplot2)
library(neuralnet)
library(iml)
library(rpart)
library(readr)
library(magrittr)
library(tibble)
library(skimr)
library(rpart.plot)
library(randomForest)
library(caret)
library(e1071)
library(DALEX)

opts_chunk$set(
collapse = TRUE,
fig.width = 6,
fig.height = 6,
fig.align = "center",
warning = FALSE,
screenshot.force = FALSE
)
```


## Studi Kasus 1: Dataset perumahan di Boston 

```{r eval=FALSE}
data("Boston", package = "MASS")
Boston
```

```{r echo=FALSE}
data("Boston", package = "MASS")
kable(head(Boston))
```

### CART

```{r}
# First we fit a machine learning model on the Boston housing data
rf <- rpart(medv ~ ., data = Boston)
```

```{r}
#| label: fig-cart-1
#| fig-cap: "Pohon keputusan 1"
#| fig-width: 4
#| fig-height: 4
#| column: margin
#| message: false
rpart.plot::rpart.plot(rf)
```


```{r}
X <- Boston[-which(names(Boston) == "medv")]
mod <- Predictor$new(rf, data = X)
```

### SHAP

```{r}
# Then we explain the first instance of the dataset with the Shapley method:
x.interest <- X[1, ]
shapley <- Shapley$new(mod, x.interest = x.interest)
shapley
```

```{r}
# Look at the results in a table
kable(shapley$results)
```


```{r}
#| fig-height: 4
# Or as a plot
plot(shapley)
```


```{r}
#| fig-height: 4
# Explain another instance
shapley$explain(X[2, ])
plot(shapley)
```


## Studi Kasus 2: Dataset Iris

Ilustrasi berikut digunakan untuk permasalahan multikelas


### CART


```{r}
rf_iris <- rpart(Species ~ ., data = iris)
```


```{r}
#| label: fig-cart-2
#| fig-cap: "Pohon keputusan 2"
#| fig-width: 4
#| fig-height: 4
#| column: margin
#| message: false
rpart.plot::rpart.plot(rf_iris)
```

```{r}
X <- iris[-which(names(iris) == "Species")]
mod <- Predictor$new(rf_iris, data = X, type = "prob")
```


### SHAP

```{r}
# Then we explain the first instance of the dataset with the Shapley() method:
shapley <- Shapley$new(mod, x.interest = X[1, ])
kable(shapley$results)
```


```{r}
#| fig-height: 4
plot(shapley)
```

```{r}
# You can also focus on one class
mod <- Predictor$new(rf_iris, data = X, type = "prob", class = "setosa")
shapley <- Shapley$new(mod, x.interest = X[1, ])
shapley$results
```


```{r}
#| fig-height: 4
plot(shapley)
```


## Studi Kasus 3: Pinjaman Bank

Suatu perusahaan perbankan meneliti 75 jenis skema pinjaman yang telah diberi rating oleh para customernya pada `loan.csv` berikut


```{r eval=FALSE}
# Read the Data
loan <- read.csv("loan.csv", header=T)
loan
```

```{r echo=FALSE}
# Read the Data
loan <- read.csv("loan.csv", header=T)
kable(head(loan))
```

Peubah yang digunakan ialah:

* Besar pinjaman (dalam juta rupiah)
* Lama pembayaran (dalam tahun)
* Tambahan bunga yang ditetapkan (dalam %)
* Pembayaran per bulan (dalam 10000)
* Banyak cash back yang diterapkan pada skema tersebut

Tujuan penelitian yang dilakukan ialah memprediksi rating skema pinjaman berdasarkan variabel-variabel tersebut. Manakah variabel yang paling penting?


### Persiapan data

```{r}
# Random sampling
samplesize <- 0.60 * nrow(loan)

set.seed(80)
index <- sample( seq_len(nrow(loan)), size = samplesize )
```


```{r}
# Create training and test set
datatrain <- loan[ index, ]
datatest <- loan[ -index, ]
```


### Penskalaan

Langkah pertama adalah penskalaan gugus data. Hal ini penting karena jika tidak, suatu peubah mungkin mempunyai dampak besar pada peubah hasil prediksi hanya karena skalanya. Terkadang peubah yang belum dilakukan proses `scaling` cenderung menghasilkan hasil yang tidak memiliki makna.

```{r}
## Scale data for neural network
max <- apply(loan , 2 , max)
min <- apply(loan, 2 , min)
scaled <- as.data.frame(scale(loan, center = min, scale = max - min))
kable(head(scaled))
```


### ANN dengan `neuralnet`

Selain terdapat di dalam paket `keras` dan `tensorflow`, komunitas R juga mengembangkan model jaringan syaraf tiruan (ANN) pada paket `neuralnet`.

#### Partisi data

```{r}
# creating training and test set
trainNN <- scaled[index , ]
testNN <- scaled[-index , ]
```

#### Membangun model

```{r}
# fit neural network
set.seed(2)
NN <- neuralnet(
  rating ~ besar.pinjaman + lama.pembayaran + bunga + pembayaran.per.bulan + banyak.cash.back, trainNN, 
  hidden = 3 , 
  linear.output = T
)
```

#### Visualisasi ANN

```{r}
# plot neural network
plot(NN)
```

#### Prediksi 

```{r}
## Prediction using neural network
predict_testNN <- compute(NN, testNN[,c(1:5)])
predict_testNN <- (predict_testNN$net.result * (max(loan$rating) - min(loan$rating))) + min(loan$rating)
```


```{r}
#| fig-height: 4
plot(datatest$rating, predict_testNN, col='blue', pch=16, ylab = "predicted rating NN", xlab = "real rating")

abline(0,1)
```


```{r}
# Calculate Root Mean Square Error (RMSE)
RMSE.NN <- (sum((datatest$rating - predict_testNN)^2) / nrow(datatest)) ^ 0.5
cat("Metrik RMSE: ", RMSE.NN)
```

### Feature importance

```{r}
#| fig-height: 4
X <- loan[which(names(loan) != "rating")]
predictor <- Predictor$new(NN, data = X, y = loan$rating)

imp <- FeatureImp$new(predictor, loss = "mae")
plot(imp)
```

```{r}
imp$results
```

### SHAP


```{r}
#| fig-height: 4
shapley <- Shapley$new(predictor, x.interest = X[1, ])
shapley$plot()
```

## Studi Kasus 4: Prediksi Gejala Stroke

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.[^3]

This dataset is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relavant information about the patient.[^4]

[^3]: [https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset](https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset)

[^4]: [https://arxiv.org/pdf/1904.11280.pdf](https://arxiv.org/pdf/1904.11280.pdf)

Attribute Information:

1. `id`: unique identifier
2. `gender`: "Male", "Female" or "Other"
3. `age`: age of the patient
4. `hypertension`: 0 if the patient doesn't have hypertension, 1 if the patient has hypertension
5. `heart_disease`: 0 if the patient doesn't have any heart diseases, 1 if the patient has a heart disease
6. `ever_married`: "No" or "Yes"
7. `work_type`: "children", "Govt_jov", "Never_worked", "Private" or "Self-employed"
8. `Residence_type`: "Rural" or "Urban"
9. `avg_glucose_level`: average glucose level in blood
10. `bmi`: body mass index
11. `smoking_status`: "formerly smoked", "never smoked", "smokes" or "Unknown"*
12. `stroke`: 1 if the patient had a stroke or 0 if not

**Note**: *Unknown* in `smoking_status` means that the information is unavailable for this patient
